# Dimensionality reduction techniques (week 5)

*This week we first continue with human data wrangling from last week, and then analysize the data using dimensionality reduction techniques.*


## 1. Data wrangling continuing from last week
*you can see it in my github:*
<https://github.com/tzsu/IODS-project/blob/master/data/create_human.R>

## 2. Analysis 

### 1. About the Human data: 
*The Human data includes two data sets: "Human development" and "Gender inequality". We keep only the countries in both data sets, and exclude rows with missing values . After combining some variables and excluding some, the **Human** data we will use for performing analysis has 155 observations and the folowing 8 variables:*

####the 8 variables:
- ratioedu2.fm = Share of female population with secondary education over(/) share of male population with secondary education
- ratiolab.fm = Share of female population that participates in labour force over(/) share of male population that participates in labour force
- exp.y.edu = Expected years of schooling
- life.exp = Life expectancy at birth
- gni.cap = Gross national income per capita
- mmr = Maternal mortality ratio
- ad.b.r = Adolescent birth rate
- par% = population with at least some secondary education


```{r, echo=TRUE}
#set wokring directory and load the human data into R

setwd("D:/Sydney Uni/forth semester units of study/Intro to Open Data Science/IODS-project")

human <- as.data.frame(read.table('data/human.csv', sep="\t", header = TRUE))
```

```{r, echo=TRUE}
#explore the structure and dimentions of the data
str(human)
```

### 2. Graphical overview and interpretation

#### From the graph below we can see that...

```{r, include=FALSE}
#access package
library(dplyr)
library(ggplot2)
library(GGally)
```

```{r, echo=TRUE}
ggpairs(human)
summary(human)
```



