# Dimensionality reduction techniques (week 5)

*This week we first continue with human data wrangling from last week, and then analysize the data using dimensionality reduction techniques.*


## 1. Data wrangling continuing from last week
*you can see it in my github:*
<https://github.com/tzsu/IODS-project/blob/master/data/create_human.R>

## 2. Analysis 

### 1. About the Human data: 
*The Human data includes two data sets: "Human development" and "Gender inequality". We keep only the countries in both data sets, and exclude rows with missing values . After combining some variables and excluding some, the **Human** data we will use for performing analysis has 155 observations and the folowing 8 variables:*

####the 8 variables:
- ratioedu2.fm = Share of female population with secondary education over(/) share of male population with secondary education
- ratiolab.fm = Share of female population that participates in labour force over(/) share of male population that participates in labour force
- exp.y.edu = Expected years of schooling
- life.exp = Life expectancy at birth
- gni.cap = Gross national income per capita
- mmr = Maternal mortality ratio
- ad.b.r = Adolescent birth rate
- par% = population with at least some secondary education


```{r, echo=TRUE}
#set wokring directory and load the human data into R

setwd("D:/Sydney Uni/forth semester units of study/Intro to Open Data Science/IODS-project")

human <- as.data.frame(read.table('data/human.csv', sep="\t", header = TRUE))
```

```{r, echo=TRUE}
#explore the structure and dimentions of the data
str(human)
```

### 2. Graphical overview and interpretation

#### From the graph below we can see that "gni.cap", "mmr", and "ad.b.r" are right skewed, whereas, "exp.y.edu" is close to normal distribution. "exp.y.edu"(expected year of education) and "life.exp"(life expectancy at birth) seems to be correlated.

```{r, include=FALSE}
#access package
library(dplyr)
library(ggplot2)
library(GGally)
# access packages FactoMineR and tidyr
library(FactoMineR)
library(tidyr)
library(corrplot)
```

```{r, echo=TRUE}
ggpairs(human)
summary(human)
```

### 3. PCA (principal component analysis) 

#### **Not standardized**: The biplot displays the observations by the first two principal components, along with arrows representing the original variables.

```{r}
#perform pca
pca_human <- prcomp(human)
#show summary of the non-standardized variables
summary(pca_human)
#draw a biplot of the principal component
```

#### **Standardized**: The means of every variables will become 0. I don't really understand this part...

```{r, echo=TRUE}
# standardize human data
human_std <- scale(human)

# print out summaries of the standardized variables
summary(human_std)

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

## 4. It's tea time

```{r, echo=TRUE}

#load tea data to R and glimpse it
data("tea")
glimpse(tea)

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# look at the summaries and structure of the data
summary(tea_time)
str(tea_time)

# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```
